---
title: "HS GCMS Spectral Deconvolution"
author: "Jefferson Pastuna"
date: "2025-03-25"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

### Introduction

Write an Introduction.

### Before to start

Write a package description.

### eRah package workflow

How to star.

```{r echo=TRUE, message=FALSE, warning=FALSE}

# eRah package installation
#install.packages('erah')
# eRah library call
library(erah)

```

Folder files

```{r echo=TRUE}

# Delete all file that are not in folders
unlink('Data/Data_to_eRah/*')
# Data folder path
createdt('Data/Data_to_eRah/')

```

New experiment.


```{r echo=TRUE}

# Loading (*.CDF) chromatograms
instrumental <- read.csv('Data/Metadata_to_eRah/HS_GCMS_Data_inst_CDF.csv')
# Loading metadata of the chromatograms (phenotype file)
phenotype <- read.csv('Data/Metadata_to_eRah/HS_GCMS_Data_pheno.csv')
# Merge of metadata information with chromatograms
ex <- newExp(instrumental = instrumental,
             phenotype = phenotype,
             info = 'Croton lechleri sap')

```

Compound deconvolution

Parameter

```{r echo=TRUE}

ex.dec.par <- setDecPar(min.peak.width = 3,
                        min.peak.height = 400,
                        noise.threshold = 40,
                        avoid.processing.mz = c(30:32))

```

parallel processing

```{r echo=TRUE}

plan(future::multisession,
     workers = 5)

```

Deconvolution

```{r echo=TRUE, warning=FALSE}

ex <- deconvolveComp(ex,
                     ex.dec.par)

```

Alignment

```{r echo=TRUE}

# Alignment parameters
ex.al.par <- setAlPar(min.spectra.cor = 0.85,
                      max.time.dist = 2.4,
                      mz.range = 32:550)
# Alignment
ex <- alignComp(ex,
                alParameters = ex.al.par)

```

Missing compound recovery


```{r echo=TRUE}

ex <- recMissComp(ex,
                  min.samples = 3)

```

Exporting alignment feature list.

```{r echo=TRUE}

# Extracting alignment feature list
feat_list <- alignList(ex,
                       by.area = FALSE)
# Exporting alignment feature list
#write.csv(feat_list,
#          file = "Result/eRah_Result/eRah_feature_table.csv")

```

# Identification

Metabolite identification was by comparing all the spectra found against a reference database. The eRah default database was used to metabolite identification. Metabolite identification was improve by exporting all the spectra found (.msp) in eRah to NIST MS Search 2.4 software.

```{r echo=TRUE, warning=FALSE}

# Identification
peak_iden <- identifyComp(ex,
                          id.database = mslib,
                          mz.range = NULL,
                          n.putative = 1)

```

Exporting spectra to NIST MS Search software identification with NIST-20 library.

```{r echo=TRUE}

#export2MSP(peak_iden,
#           store.path = "Result/eRah_Result",
#           alg.version = 2)

```

